SPECS-decadal-MOS

### 2017-04-28 ClimaTrans/climatrans.Rmd
ClimaTrans

@@

```{r,echo=TRUE}
library(esd)
library(ncdf)
setwd('~/R')
#rm(list=ls())
## Information about the system and session
```

@@

#### Pre-processing: adapting the netCDF data to esd: set up station objects: 
```{r,eval=FALSE}
## Precipitation
## time origin: minutes since 1901-01-01 00:00
library(ncdf)
ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/delhi.nc')
delhi <- get.var.ncdf(ncid,varid='p')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)
delhi <- zoo(x=delhi,order.by=as.Date(time/(24*60),origin='1901-01-01'))
delhi <- as.station(delhi,loc='Delhi',param='precip',unit='mm/day',lon=77.2,lat=28.6,
                    cntr='India',longname='precipitation',
                    info='extracted rainfall data from IMD gridded daily data',
                    ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bombay.nc')
bombay <- get.var.ncdf(ncid,varid='p')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)
bombay <- zoo(x=bombay,order.by=as.Date(time/(24*60),origin='1901-01-01'))
bombay <- as.station(bombay,loc='Bombay',param='precip',unit='mm/day',lon=72.9,lat=19.0,
                cntr='India',longname='precipitation',
                info='extracted rainfall data from IMD gridded daily data',
                ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bangalore.nc')
bangalore <- get.var.ncdf(ncid,varid='p')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)

@@

## Maximum temperature
## time origin: minutes since 1969-01-01 00:00
ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/delhi_tmax.nc')
delhi <- get.var.ncdf(ncid,varid='temp')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)
delhi <- zoo(x=delhi,order.by=as.Date(time/(24*60),origin='1969-01-01'))
tx.delhi <- as.station(delhi,loc='Delhi',param='tmax',unit='degC',lon=77.2,lat=28.6,
               cntr='India',longname='daily maximum temperature',
               info='extracted from IMD gridded daily data',
               ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bombay_tmax.nc')
bombay <- get.var.ncdf(ncid,varid='temp')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)
bombay <- zoo(x=bombay,order.by=as.Date(time/(24*60),origin='1969-01-01'))
tx.bombay <- as.station(bombay,loc='Bombay',param='tmax',unit='degC',lon=72.9,lat=19.0,
                cntr='India',longname='daily maximum temperature',
                info='extracted from IMD gridded daily data',
                ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- open.ncdf('~/Dropbox/Public/ClimaTrans/bangalore_tmax.nc')
bangalore <- get.var.ncdf(ncid,varid='temp')
time <- get.var.ncdf(ncid,varid='time')
close.ncdf(ncid)
bangalore <- zoo(x=bangalore,order.by=as.Date(time/(24*60),origin='1969-01-01'))
tx.bangalore <- as.station(bangalore,loc='Bangalore',param='tmax',unit='degC',
                lon=77.6,lat=13.0,cntr='India',longname='daily maximum temperature',
                info='extracted from IMD gridded daily data',
                ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

@@

```{r, fig.width=6, fig.height=6}
## Wet-day frequency
plot(fw,new=FALSE)
for(i in 1:3) lines(trend(subset(fw,is=i)))
grid()
print(trend(fw,result='pval')) 
```

@@

```{r, fig.width=6, fig.height=6}
## Wet-day mean
plot(mu,new=FALSE)
for(i in 1:3) lines(trend(subset(mu,is=i)))
grid()
print(trend(mu,result='pval'))
```

@@

```{r, fig.width=6, fig.height=6}
## Probability of exceedin 50 mm/day
Pr <- PrexpPr(mu,fw,x0=50) 
plot(Pr,new=FALSE,errorbar=FALSE)
for(i in 1:3) lines(trend(subset(Pr,is=i)))
grid()
print(trend(Pr,result='pval'))
```

@@

```{r, fig.width=6, fig.height=6}
## Aggregate to annual totals
pt <- annual(climatrans.pr,'sum')
attr(pt,'unit') <- 'mm/season'
plot(pt,new=FALSE)
for(i in 1:3) lines(trend(subset(pt,is=i)))
grid()
print(trend(pt,result='pval'))
```

@@

```{r}
## Correlation between the wet-day frequency and the SLP
z <- subset(fw,is=2)
corfield(z,slp,colbar=list(pal='t2m',breaks=seq(-1,1,by=0.1)),new=FALSE)
```

@@

```{r}
ds.fw <- DS(pca.fw,eof.slp,eofs=1:20)
plot(ds.fw,new=FALSE)
```

@@

```{r pca-of-residual-fw}
## Examine the PCA of residuals for the cold monthsto see if there are any structures left
  fw.2 <- as.station(as.residual(ds.fw))
  pca.fw.2 <- PCA(fw.2)
  plot(pca.fw.2,new=FALSE)
  grid()
```

@@

```{r}
  ds.tx <- DS(pca.tx,eof.t2m)
  plot(ds.tx,new=FALSE)
```

@@

```{r}
## Check the residuals
  tx.2 <- as.station(as.residual(ds.tx))
  pca.tx.2 <- PCA(tx.2)
  plot(pca.tx.2,new=FALSE)
```

@@

```{r fig.width=12, fig.height=20}
print('Compare common EOFs')
diag.slp <- biasdiag(slp,it=month.abb[season],verbose=FALSE)
```

>>>>

To extract just the R-code:

```{r, eval=FALSE}
library(knitr)
purl('~/git/esd_Rmarkdown/ClimaTrans/climatrans.Rmd', output='~/git/esd_Rmarkdown/ClimaTrans/climatrans.R')
```

@@

```{r,echo=TRUE}
library(esd)
library(ncdf4)
setwd('~/R')
#rm(list=ls())
## Information about the system and session
```

@@

#### Pre-processing: adapting the netCDF data to esd: set up station objects: 
```{r,eval=FALSE}
## Precipitation
## time origin: minutes since 1901-01-01 00:00
library(ncdf4)
ncid <- nc_open('~/Dropbox/Public/ClimaTrans/delhi.nc')
delhi <- ncvar_get(ncid,varid='p')
time <- ncvar_get(ncid,varid='time')
nc_close(ncid)
delhi <- zoo(x=delhi,order.by=as.Date(time/(24*60),origin='1901-01-01'))
delhi <- as.station(delhi,loc='Delhi',param='precip',unit='mm/day',lon=77.2,lat=28.6,
                    cntr='India',longname='precipitation',
                    info='extracted rainfall data from IMD gridded daily data',
                    ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- nc_open('~/Dropbox/Public/ClimaTrans/bombay.nc')
bombay <- ncvar_get(ncid,varid='p')
time <- ncvar_get(ncid,varid='time')
nc_close(ncid)
bombay <- zoo(x=bombay,order.by=as.Date(time/(24*60),origin='1901-01-01'))
bombay <- as.station(bombay,loc='Bombay',param='precip',unit='mm/day',lon=72.9,lat=19.0,
                cntr='India',longname='precipitation',
                info='extracted rainfall data from IMD gridded daily data',
                ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- nc_open('~/Dropbox/Public/ClimaTrans/bangalore.nc')
bangalore <- ncvar_get(ncid,varid='p')
time <- ncvar_get(ncid,varid='time')
nc_close(ncid)

@@

## Maximum temperature
## time origin: minutes since 1969-01-01 00:00
ncid <- nc_open('~/Dropbox/Public/ClimaTrans/delhi_tmax.nc')
delhi <- ncvar_get(ncid,varid='temp')
time <- ncvar_get(ncid,varid='time')
nc_close(ncid)
delhi <- zoo(x=delhi,order.by=as.Date(time/(24*60),origin='1969-01-01'))
tx.delhi <- as.station(delhi,loc='Delhi',param='tmax',unit='degC',lon=77.2,lat=28.6,
               cntr='India',longname='daily maximum temperature',
               info='extracted from IMD gridded daily data',
               ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- nc_open('~/Dropbox/Public/ClimaTrans/bombay_tmax.nc')
bombay <- ncvar_get(ncid,varid='temp')
time <- ncvar_get(ncid,varid='time')
nc_close(ncid)
bombay <- zoo(x=bombay,order.by=as.Date(time/(24*60),origin='1969-01-01'))
tx.bombay <- as.station(bombay,loc='Bombay',param='tmax',unit='degC',lon=72.9,lat=19.0,
                cntr='India',longname='daily maximum temperature',
                info='extracted from IMD gridded daily data',
                ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

ncid <- nc_open('~/Dropbox/Public/ClimaTrans/bangalore_tmax.nc')
bangalore <- ncvar_get(ncid,varid='temp')
time <- ncvar_get(ncid,varid='time')
nc_close(ncid)
bangalore <- zoo(x=bangalore,order.by=as.Date(time/(24*60),origin='1969-01-01'))
tx.bangalore <- as.station(bangalore,loc='Bangalore',param='tmax',unit='degC',
                lon=77.6,lat=13.0,cntr='India',longname='daily maximum temperature',
                info='extracted from IMD gridded daily data',
                ref='Madhusoodanan M.S. ("Madhusoodanan M.S." <madhusoodanan@gmail.com>)')

@@

```{r, fig.width=6, fig.height=6}
## Wet-day frequency
plot(fw,new=FALSE)
for(i in 1:3) lines(trend(subset(fw,is=i)))
grid()
print(as.numeric(trend(fw,result='coef'))) 
print(as.numeric(trend(fw,result='pval'))) 
```

@@

```{r, fig.width=6, fig.height=6}
## Wet-day mean
plot(mu,new=FALSE)
for(i in 1:3) lines(trend(subset(mu,is=i)))
grid()
print(as.numeric(trend(mu,result='coef'))) 
print(as.numeric(trend(mu,result='pval')))
```

@@

```{r, fig.width=6, fig.height=6}
## Probability of exceedin 50 mm/day
Pr <- PrexpPr(mu,fw,x0=50) 
plot(Pr,new=FALSE,errorbar=FALSE)
for(i in 1:3) lines(trend(subset(Pr,is=i)))
grid()
print(as.numeric(trend(Pr,result='coef')))
print(as.numeric(trend(Pr,result='pval')))
```

@@

```{r, fig.width=6, fig.height=6}
## Aggregate to annual totals
pt <- annual(climatrans.pr,'sum')
attr(pt,'unit') <- 'mm/season'
plot(pt,new=FALSE)
for(i in 1:3) lines(trend(subset(pt,is=i)))
grid()
print(as.numeric(trend(pt,result='pval')))
print(as.numeric(trend(pt,result='pval')))
```

@@

```{r}
## Correlation between the wet-day frequency and the SLP
z <- subset(fw,is=2)
attr(z,'unit') <- 'correlation'
corfield(z,slp,colbar=list(pal='t2m',breaks=seq(-1,1,by=0.1)),new=FALSE)
```

@@

```{r, fig.height=8,fig.width=8}
ds.fw <- DS(pca.fw,eof.slp,eofs=1:20)
plot(ds.fw,new=FALSE)
```

@@

```{r pca-of-residual-fw, fig.height=8,fig.width=8}
## Examine the PCA of residuals for the cold monthsto see if there are any structures left
  fw.2 <- as.station(as.residual(ds.fw))
  pca.fw.2 <- PCA(fw.2)
  plot(pca.fw.2,new=FALSE)
  grid()
```

@@

```{r, fig.height=8,fig.width=8}
  ds.tx <- DS(pca.tx,eof.t2m)
  plot(ds.tx,new=FALSE)
```

@@

```{r, fig.height=8,fig.width=8}
## Check the residuals
  tx.2 <- as.station(as.residual(ds.tx))
  pca.tx.2 <- PCA(tx.2)
  plot(pca.tx.2,new=FALSE)
```

@@

```{r fig.width=12, fig.height=20}
print('Compare common EOFs')
diag.slp <- biasdiag(slp,it=month.abb[season],verbose=FALSE)
mean.diff <- unlist(lapply(diag.slp,function(x) x$mean.diff[,1]))
sd.ratio <- 1 - unlist(lapply(diag.slp,function(x) x$sd.ratio[,1]))
ar.ratio <- 1 - unlist(lapply(diag.slp,function(x) x$autocorr.ratio[,1]))
summary(sd.ratio)
summary(ar.ratio)
```

### 2017-04-28 EU-Circle/eu-circle_case-study-1.Rmd
ClimaTrans

@@

```{r,warning=FALSE}
X <- retrieve('data.ECAD/rr_0.25deg_reg.nc',lon=c(4,8),lat=c(43,46))
map(X,new=FALSE)
```

@@

>>>>

```{r, eval=FALSE}
library(knitr)
purl('~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.Rmd', output='~/git/esd_Rmarkdown/EU-Circle/eu-circle_case-study-1.R')
```

@@

```{r,warning=FALSE}
X <- retrieve('~/data.ECAD/rr_0.25deg_reg.nc',lon=c(4,8),lat=c(43,46))
map(X,new=FALSE)
```

@@

Calculate EOFs which can serve as predictors for the downscaling.

```{r, fig.height=8}
dryseason <- function(x,na.rm=TRUE) if (sum(is.finite(x))>1) mean(subset(spell(x,threshold=1),is=2),na.rm=na.rm) else NA
if (!file.exists('case1-dryseason.rda')) {
  z <- aggregate(subset(X,it='mjjas'),year,FUN='dryseason')
  save(z,file='case1-dryseason.rda')
} else load('case1-dryseason.rda')
eof.dry <- EOF(z)
plot(eof.dry)
```

Predictors for calibration: check for any dependency

```{r}
T2M <- aggregate(subset(retrieve('~/Downloads/air.mon.mean.nc',lon=c(-12,35),lat=c(27,50)),it='mjjas'),year,'mean')
SLP <- aggregate(subset(retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-12,35),lat=c(27,50)),it='mjjas'),year,'mean')
eof.t2m <- EOF(T2M)
eof.slp <- EOF(SLP)
```

Any connection between temperature and dry spells?

```{r, fig.height=8}
ds.t2m <- DS(eof.dry,eof.t2m)
plot(ds.t2m)
```

Any connection between sea-level pressure anomalies and dry spells?

```{r, fig.height=8}
ds.slp <- DS(eof.dry,eof.slp)
plot(ds.slp)
```

### 2017-04-28 SPECS/specs-decade-mos.Rmd
ClimaTrans

  ## Apply an EOF-filter to emphasise only the large scale anomalies:
  if (!is.null(eofs)) {
    print(paste('Only keep the large-scale structures; EOFs:',paste(eofs,collapse=' ,')))
    Xm.eofs <- subset(EOF(Xm),pattern=eofs)
    Xm <- eof2field(Xm.eofs)
    land <- paste(land,'eof_',paste(eofs,collapse='+'))
  }

>>>>

  ## Apply an EOF-filter to emphasise only the large scale anomalies:
  if (!is.null(eofs)) {
    print(paste('Only keep the large-scale structures; EOFs:',paste(eofs,collapse=' ,')))
    Xm.eofs <- subset(EOF(Xm),ip=eofs)
    Xm <- eof2field(Xm.eofs)
    land <- paste(land,'eof_',paste(eofs,collapse='+'))
  }
